---
title: "Example Gallery"
output: 
  rmarkdown::html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Most of these examples are adapted from the [Vega-Lite example gallery](https://vega.github.io/vega-lite/examples/).See that example gallery for more examples of plots that can be made with Vega-Lite and thus likely can be built with `vlbuildr`.

```{r}
library (vlbuildr)
```


## Single-view plots

### Bar charts {.tabset}

#### Simple Bar Chart

```{r}
values <- tibble::tribble(
  ~a, ~b,
  "A", 28, 
  "B", 55, 
  "C",  43,
  "D", 91, 
  "E",  81, 
  "F", 53,
  "G",19, 
  "H", 87, 
  "I", 52
)

vl_chart(values) %>% 
  vl_mark_bar() %>%
  vl_encode_x('a') %>% 
  vl_axis_x(labelAngle = 0) %>%
  vl_encode_y('b')
```

#### Aggregate Bar Chart 

```{r aggregate-bar-chart}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/population.json") %>%
  vl_filter("datum.year == 2000") %>%
  vl_encode_y(field = "age", 
              type = "ordinal") %>%
  vl_encode_x(field = "people", 
              type = "quantitative", 
              aggregate = "sum",
              axis = list(title = "population"), 
              stack = "normalize") %>%
  vl_mark_bar() %>%
  vl_sort_y_by_encoding(encoding = "x")

```

#### Stacked Bar Chart

```{r stacked-bar-chart}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/seattle-weather.csv") %>%
  vl_mark_bar() %>%
  vl_encode_x(field = 'date', type = 'ordinal', timeUnit = 'month', 
              title = 'Month of the year') %>%
  vl_encode_y(aggregate = 'count', type = 'quantitative') %>%
  vl_encode_color('weather', type = 'nominal', title = 'Weather type') %>%
  vl_scale_color(domain = list("sun", "fog", "drizzle", "rain", "snow"),
                 range = list("#e7ba52", "#c7c7c7", "#aec7e8", "#1f77b4", "#9467bd"))
```

#### Normalized Stacked Bar Chart 

```{r normalized-stacked-bar-chart}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/population.json") %>%
  vl_calculate(calculate = "datum.sex == 2 ? 'Female' : 'Male'", 
               as = "gender") %>%
  vl_filter("datum.year == 2000") %>%
  vl_encode_x(field = "age", 
              type = "ordinal") %>%
  vl_encode_y(field = "people", type = "quantitative") %>%
  vl_stack_y("normalize") %>%
  vl_aggregate_y(aggregate = "sum") %>%
  vl_axis_y(title = "population") %>%
  vl_encode_color(field = "gender", type = "nominal") %>%
  vl_mark_bar() %>% 
  vl_config_axis(domainWidth = 1) %>%
  vl_config_view(stroke = "transparent")
```




### Histograms and Density Plots {.tabset}

#### Histogram 

```{r histogram}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/movies.json") %>%
  vl_encode_x(field = "IMDB_Rating", type = "quantitative") %>%
  vl_encode_y(type = "quantitative", aggregate = "count") %>%
  vl_bin_x(maxbins=10) %>%
  vl_mark_bar()
```

#### 2D Histogram

```{r 2d-histogram}
vl_chart() %>%
  vl_add_data(url = 'https://vega.github.io/vega-editor/app/data/movies.json') %>%
  vl_filter('isDefined(datum["IMDB_Rating"]) && isDefined(datum["Rotten_Tomatoes_Rating"])') %>%
  vl_encode_x(field = 'IMDB_Rating', type = 'quantitative') %>%
  vl_encode_y(field = 'Rotten_Tomatoes_Rating', type = 'quantitative') %>%
  vl_encode_color(type = 'quantitative', aggregate = 'count') %>%
  vl_bin_x(maxbins=60) %>%
  vl_bin_y(maxbins=40) %>%
  vl_mark_rect() %>%
  vl_config_view(stroke = 'transparent')
```

#### Density Plot

```{r density}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/movies.json") %>%
  vl_density("IMDB_Rating", bandwidth = 0.3) %>%
  vl_encode_x(field = "value", type = "quantitative", title = "IMDB Rating") %>%
  vl_encode_y(field = "density", type = "quantitative") %>%
  vl_mark_area()
```


### Scatter & Strip Plots {.tabset}

#### Scatterplot

```{r scatter}
vl_chart() %>%
   vl_add_data(values = mtcars) %>%
   vl_mark_point() %>%
   vl_encode_x("wt") %>%
   vl_encode_y("mpg") %>% 
   vl_encode_tooltip_array(c("_row","mpg","wt"))
```


#### 1D Striplot

```{r 1d-strip}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/seattle-weather.csv") %>%
  vl_encode_x(field = "precipitation", type = "quantitative") %>%
  vl_mark_tick()
```


#### Strip plot

```{r strip}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_encode_x(field = "Horsepower", type = "quantitative") %>%
  vl_encode_y(field = "Cylinders", type = "ordinal") %>%
  vl_mark_tick()
```

#### Bubble plot

```{r bubble-gapminder}
vl_chart(width = 500, height = 300) %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/gapminder-health-income.csv") %>%
  vl_add_interval_selection("view", bind = "scales") %>%
  vl_mark_circle() %>%
  vl_encode_x(field = 'income', type = "quantitative") %>%
  vl_scale_x(type = 'log') %>%
  vl_encode_y(field = 'health', type = "quantitative") %>%
  vl_scale_y(zero = FALSE) %>%
  vl_axis_y(minExtent = 30) %>%
  vl_encode_size(field = "population", type = 'quantitative') %>%
  vl_encode_color(value = '#000')
```

#### Jittered scatter plot

```{r}
vl_chart(height = list(step = 50)) %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_calculate("random()", as = "random") %>%
  vl_mark_point() %>%
  vl_encode_x("Horsepower:Q") %>%
  vl_encode_y("Cylinders:O") %>%
  vl_encode_yOffset("random:Q") %>%
  vl_scale_yOffset(padding = 5) %>% 
  vl_encode_color("Cylinders:N")
```


### Line Charts {.tabset}

#### Line Chart

```{r}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/stocks.csv") %>%
  vl_filter('datum.symbol==="GOOG"') %>%
  vl_mark_line() %>%
  vl_encode_x('date:T') %>%
  vl_encode_y('price:Q')
```

#### Multi-series line chart with point markers

```{r}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/stocks.csv") %>%
  vl_mark_line(point = TRUE) %>%
  vl_encode_x('date', timeUnit = 'year') %>%
  vl_encode_y('price:Q', aggregate = 'mean') %>%
  vl_encode_color('symbol:N')
```

#### Multi-series line chart with repeat operator

```{r multi-series-line-repeat}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/movies.json") %>%
  vl_repeat_layer('US_Gross', 'Worldwide_Gross') %>%
  vl_encode_x(field = "IMDB_Rating", type = "quantitative", bin = TRUE) %>%
  vl_encode_y(field = list(`repeat` = "layer"), type = "quantitative", aggregate = "mean", 
              title = "Mean of US and Worldwide Gross") %>% 
  vl_encode_color(datum = list(`repeat` = "layer"), type = "nominal") %>%
  vl_mark_line()

```

#### Connected Scatterplot

```{r connected-scatterplot}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/driving.json") %>%
  vl_encode_x('miles:Q') %>%
  vl_scale_x(zero = FALSE) %>%
  vl_encode_y('gas:Q') %>%
  vl_scale_y(zero = FALSE) %>%
  vl_encode_order('year') %>%
  vl_mark_line(point = TRUE)

```

#### Slope Graph

```{r slope-graph}
vl_chart(width = list(step = 50)) %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/barley.json") %>%
  vl_mark_line() %>%
  vl_encode_x('year:O') %>%
  vl_scale_x(padding = 0.5) %>%
  vl_encode_y('yield:Q', aggregate = 'median') %>%
  vl_encode_color('site:N') %>%
  vl_mark_line()

```

### Area charts & streamgraphs {.tabset}

#### Area chart

```{r area}
vl_chart(data = 'https://vega.github.io/vega-editor/app/data/unemployment-across-industries.json', width = 300, height = 200) %>%
  vl_mark_area() %>%
  vl_encode_x(field = 'date', timeUnit = 'yearmonth') %>%
  vl_axis_x(format = '%Y') %>% 
  vl_encode_y(field = 'count', aggregate = 'sum') 
```

#### Stacked Area chart

```{r stacked-area}
vl_chart(data = 'https://vega.github.io/vega-editor/app/data/unemployment-across-industries.json', width = 300, height = 200) %>%
  vl_mark_area() %>%
  vl_encode_x(field = 'date', timeUnit = 'yearmonth') %>%
  vl_axis_x(format = '%Y') %>% 
  vl_encode_y(field = 'count', aggregate = 'sum') %>%
  vl_encode_color('series') %>%
  vl_scale_color(scheme = 'category20b')
```

#### Normalized stacked aea chart

```{r normalized-area}
vl_chart(data = 'https://vega.github.io/vega-editor/app/data/unemployment-across-industries.json', width = 300, height = 200) %>%
  vl_mark_area() %>%
  vl_encode_x(field = 'date', timeUnit = 'yearmonth') %>%
  vl_axis_x(format = '%Y') %>% 
  vl_encode_y(field = 'count', aggregate = 'sum', stack = 'normalize', axis = NA) %>%
  vl_encode_color('series') %>%
  vl_scale_color(scheme = 'category20b')
```

#### Streamgraph

```{r stream-graph}
vl_chart(data = 'https://vega.github.io/vega-editor/app/data/unemployment-across-industries.json') %>%
  vl_mark_area() %>%
  vl_encode_x(field = 'date', timeUnit = 'yearmonth') %>%
  vl_axis_x(domain = FALSE, format = '%Y', tickSize = 0) %>% 
  vl_encode_y(field = 'count', aggregate = 'sum', stack = 'center', axis = NULL) %>%
  vl_encode_color(field = 'series') %>%
  vl_scale_color(scheme = 'category20b') %>%
  vl_encode_opacity(value = 0.2) %>%
  vl_condition_opacity(param = 'industry', value = 1) %>%
  vl_add_point_selection(name = 'industry', fields = list('series'), bind = 'legend') 
```



### Table-based plots {.tabset}

#### Heatmap

```{r weather-heatmap}

vl_chart(title = "Daily Max Temperatures (C) in Seattle, WA") %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/seattle-weather.csv") %>%
  vl_mark_rect() %>%
  vl_encode_x(field = 'date', type = 'ordinal', timeUnit = 'date', 
              title = 'Day') %>%
  vl_axis_x(labelAngle = 0, format = '%e') %>%
  vl_encode_y(field = 'date', type = 'ordinal', timeUnit = 'month', title = 'Month') %>%
  vl_encode_color('temp_max', type = 'quantitative', aggregate = 'max') %>%
  vl_legend_color(title = NA) 


```


#### Heatmap with labels

```{r heatmap-labels}
color_layer <- vl_chart() %>%
  vl_mark_rect() %>%
  vl_encode_color('num_cars:Q', title = 'Count of Records') %>%
  vl_legend_color(direction = 'horizontal')
  
text_layer <- vl_chart() %>%
  vl_mark_text() %>%
  vl_encode_text('num_cars:N') %>%
  vl_encode_color(value = "white") %>%
  vl_condition_color(test = 'datum["num_cars"] < 40', value = 'black') 
  

vl_layer(color_layer, text_layer) %>% 
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_aggregate(aggregate = list(list(op = 'count', as = 'num_cars')),
               groupby= list('Origin','Cylinders')) %>%
  vl_encode_y(field = "Origin", type = "ordinal") %>%
  vl_encode_x(field = "Cylinders", type = "ordinal") %>%
  vl_config_axis(grid = TRUE, tickBand = 'extent')
```



#### Time-series Heatmap (Lasagna Plot)

Note: This is an example where a good bit of manual nested json still needs to be made to get 
the axis sub-element conditioning.

```{r}
vl_chart(width = 300, height = 100) %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/stocks.csv") %>%
  vl_filter('datum.symbol != "GOOG"') %>%
  vl_mark_rect() %>%
  vl_encode_x('date:O', timeUnit = 'yearmonthdate', title = 'Time') %>%
  vl_axis_x(format = '%Y', labelAngle = 0, labelOverlap = FALSE,
            labelColor = list(
              condition = list( 
                test = list(
                  timeUnit = 'monthdate', 
                  field = 'value', 
                  equal = list(month = 1, date = 1)
                  ),
                value = 'black'
                ),
              value = NA
              ),
            tickColor = list(
              condition = list( 
                test = list(
                  timeUnit = 'monthdate', 
                  field = 'value', 
                  equal = list(month = 1, date = 1)
                  ),
                value = 'black'
                ),
              value = NA
              )
            ) %>%
  vl_encode_color('price:Q', aggregate = 'sum', title = 'Price') %>%
  vl_encode_y('symbol:N', title = NA)
```

### Circular plots {.tabset}

#### Pie chart 

```{r}
vl_chart(mtcars) %>%
  vl_mark_arc() %>%
  vl_encode_angle( aggregate = 'count') %>%
  vl_encode_color('cyl:N')
```

#### Donut Chart

```{r}
vl_chart(mtcars) %>%
  vl_mark_arc(innerRadius = 50) %>%
  vl_encode_angle( aggregate = 'count') %>%
  vl_encode_color('cyl:N')
```

#### Pie chart with labels

```{r}
pie_layer <- vl_chart() %>% 
  vl_mark_arc(outerRadius = 80)  %>%
  vl_encode_color('cyl:N')

text_layer <- vl_chart() %>% 
  vl_mark_text(radius = 90) %>%
  vl_encode_text('cyl:N')
 
vl_layer(pie_layer, text_layer) %>%
  vl_add_data(mtcars) %>%
  vl_encode_theta( aggregate = 'count', stack = TRUE)
```




### Advanced Calculations {.tabset}

#### Rolling average

```{r}
points <- vl_chart() %>% 
  vl_mark_point() %>%
  vl_encode_y(field = 'temp_max', 
              title = 'Max Temperature')

line <- vl_chart() %>% 
  vl_mark_line(color = 'red', size = 3) %>%
  vl_encode_y(field = 'rolling_mean', 
              title = 'Rolling Mean of Max Temperature')

vl_layer(points, line) %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/seattle-weather.csv") %>%
  vl_window(window = list(vl$Window(op = "mean", field = "temp_max", as = "rolling_mean")), frame = list(15,15)) %>%
  vl_encode_x('date:T', title = 'Date') %>%
  vl_encode_y(type = 'quantitative') %>%
  vl_axis_y(title = 'Max Temperature and Rolling Mean') %>%
  vl_add_properties(width = 400, height = 300)
```

#### Loess transform

```{r}
points <- vl_chart() %>% 
  vl_mark_point(filled = TRUE) %>%
  vl_encode_y('IMDB_Rating:Q') 

line <- vl_chart() %>% 
  vl_loess(loess = 'IMDB_Rating', on = 'Rotten_Tomatoes_Rating') %>%
  vl_mark_line(color = 'firebrick') %>%
  vl_encode_y('IMDB_Rating:Q') 

vl_layer(points, line) %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/movies.json") %>%
  vl_encode_x('Rotten_Tomatoes_Rating:Q')
```

### Composite marks: Boxplots, error bars, and error bands {.tabset}

#### Box plot

```{r}
vl_chart(data = 'https://vega.github.io/editor/data/penguins.json') %>%
  vl_mark_boxplot() %>%
  vl_encode_x('Species:N') %>%
  vl_encode_y('Body Mass (g)', type = 'quantitative') %>%
  vl_scale_y(zero = FALSE) %>%
  vl_encode_color('Species:N', legend = NA)
  
```

#### Error bars with confidence interval

```{r error-bar-ci}
points <- vl_chart() %>%
  vl_mark_point() %>%
  vl_encode_x('yield:Q', aggregate = 'mean', title = 'Barley Yield') %>% 
  vl_scale_x(zero = FALSE) %>%
  vl_encode_color(value = 'black')
  
ci <- vl_chart() %>%
  vl_mark_errorbar(extent = 'ci') %>%
  vl_encode_x('yield:Q', title = 'Barley Yield') 

vl_layer(points, ci) %>%
  vl_add_data(url = 'https://vega.github.io/vega-editor/app/data/barley.json') %>%
  vl_encode_y('variety:O') 
```

#### Error bars with standard deviation

```{r error-bar-sd}
points <- vl_chart() %>%
  vl_mark_point() %>%
  vl_encode_x('yield:Q', aggregate = 'mean', title = 'Barley Yield') %>% 
  vl_scale_x(zero = FALSE) %>%
  vl_encode_color(value = 'black')
  
ci <- vl_chart() %>%
  vl_mark_errorbar(extent = 'stdev') %>%
  vl_encode_x('yield:Q', title = 'Barley Yield') 

vl_layer(points, ci) %>%
  vl_add_data(url = 'https://vega.github.io/vega-editor/app/data/barley.json') %>%
  vl_encode_y('variety:O') 
```
#### Line chart with confidence interval band

```{r error-band}
line <- vl_chart() %>%
  vl_mark_line() %>%
  vl_encode_y('Miles_per_Gallon:Q', aggregate = 'mean')
  
ci <- vl_chart() %>%
  vl_mark_errorband(extent = 'ci') %>%
  vl_encode_y('Miles_per_Gallon:Q', title = 'Mean of Miles per Gallon (95% CIs)')

vl_layer(line, ci) %>%
  vl_add_data(url = 'https://vega.github.io/vega-editor/app/data/cars.json') %>%
  vl_encode_x('Year', timeUnit = 'year') 
```

### Maps & Geographic Displays { .tabset}

#### Chloropleth

```{r chloropleth}
vl_chart(width = 500, height = 300) %>%
  vl_add_data(
    url = 'https://vega.github.io/vega-editor/app/data/us-10m.json',
    format = list(type = 'topojson', feature = 'counties')
  ) %>%
  vl_lookup(
    lookup = 'id', 
    from = list(
      data = list(url = 'https://vega.github.io/vega-editor/app/data/unemployment.tsv'),
      key = 'id', 
      fields = list('rate')
    )
  ) %>%
  vl_add_properties(projection = list(type = 'albersUsa')) %>%
  vl_mark_geoshape() %>%
  vl_encode_color('rate:Q')
```
#### Points overlayed on U.S. map

```{r dots-map}
map <- vl_chart(projection = list(type = 'albersUsa')) %>%
  vl_add_data(
    url = 'https://vega.github.io/vega-editor/app/data/us-10m.json',
    format = list(type = 'topojson', feature = 'states')
  ) %>%
  vl_mark_geoshape(fill = 'lightgray', stroke = 'white') 

points <- vl_chart(projection = list(type = 'albersUsa')) %>%
  vl_add_data(
    url = 'https://vega.github.io/vega-editor/app/data/airports.csv',
  ) %>%
  vl_mark_circle() %>%
  vl_encode_longitude('longitude:Q') %>%
  vl_encode_latitude('latitude:Q') %>%
  vl_encode_size(value = 10) %>%
  vl_encode_color(value = 'steelblue')

vl_layer(map,points) %>%
  vl_add_properties(width = 500, height = 300)

```


### Interactive {.tabset}

#### Brush-linked color

```{r}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_add_interval_selection(name = "brush", value = list(x = c(55,160), y = c(13,37))) %>%
  vl_mark_point() %>%
  vl_encode_x(field = "Horsepower", type = "quantitative") %>%
  vl_encode_y(field = "Miles_per_Gallon", type = "quantitative") %>%
  vl_encode_color(value = "grey") %>%
  vl_condition_color(param = "brush", field = "Cylinders", type = "ordinal")
```

#### Drop-down controlled parameter

```{r}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/movies.json") %>%
  vl_encode_x(field = "IMDB_Rating", type = "quantitative") %>%
  vl_encode_y(type = "quantitative", aggregate = "count") %>%
  vl_bin_x(maxbins=10) %>%
  vl_mark_bar(cornerRadius = list(expr = "cornerRadius")) %>%
  vl_add_parameter("cornerRadius", value = 0) %>% 
  vl_bind_range_input("cornerRadius", min = 0, max = 10, step = 1)
```

#### Selection sliders

```{r}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_calculate(calculate = "year(datum.Year)", as = "Year") %>%
  vl_add_point_selection("CylYr",  fields = c("Cylinders","Year")) %>%
  vl_bind_range_input("CylYr", "Cylinders", min = 3, max = 8, step = 1) %>%
  vl_bind_range_input("CylYr", "Year", min = 1969, max = 1981, step = 1) %>%
  vl_mark_circle() %>%
  vl_encode_x(field = "Horsepower", type = "quantitative") %>%
  vl_encode_y(field = "Miles_per_Gallon", type = "quantitative") %>%
  vl_encode_color(value = "grey") %>%
  vl_condition_color(param = "CylYr", field = "Origin", type = "nominal") %>%
  vl_encode_size(value = 50) %>%
  vl_condition_size(param = "CylYr", value = 100)
```

#### Selection dropdown

```{r dropdown-select}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_add_point_selection("org",fields = list("Origin")) %>%
  vl_bind_select_input("org", options = list(NA,"Europe","Japan","USA")) %>%
  vl_mark_point() %>%
  vl_encode_x(field = "Horsepower", type = "quantitative") %>%
  vl_encode_y(field = "Miles_per_Gallon", type = "quantitative") %>%
  vl_encode_color(value = "grey") %>%
  vl_condition_color(param = "org", field = "Cylinders", type = "ordinal")
```


## Multi-view display

### Faceting / Small Multiple {.tabset}

#### Row Facet

```{r facet-row}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_mark_bar() %>%
  vl_encode_x(field = "Horsepower", type = "quantitative") %>%
  vl_bin_x(maxbins = 15) %>%
  vl_encode_y(aggregate = "count", type = "quantitative") %>%
  vl_encode_row(field = "Origin", type = "nominal")

```

#### Column Facet

```{r facet-column}
vl_chart() %>%
  vl_add_data(url = 'https://vega.github.io/vega-editor/app/data/barley.json') %>%
  vl_mark_bar() %>%
  vl_encode_y('variety:N') %>%
  vl_encode_x('yield:Q', aggregate = "sum") %>%
  vl_encode_color('site:N') %>%
  vl_encode_column("year:N")

```



#### Facet Wrap

```{r facet-wrap}
vl_chart() %>%
  vl_add_data(url = 'https://vega.github.io/vega-editor/app/data/barley.json') %>%
  vl_mark_point() %>%
  vl_encode_y('variety:N', sort = '-x') %>%
  vl_encode_x('yield:Q', aggregate = "median") %>%
  vl_scale_x(zero = FALSE) %>%
  vl_encode_color('year:N') %>%
  vl_encode_facet("site:O", columns = 2) %>%
  vl_sort_facet_by_field(op = 'median', field = 'yield')

```


#### Facetted layered plot

```{r facet-layers}
heatmap <- vl_chart() %>%
  vl_encode_color(aggregate = "count", type = "Q") %>%
  vl_scale_color(scheme = "darkred") %>%
  vl_encode_x(field = "Cylinders:O") %>%
  vl_scale_x(paddingInner = 0) %>%
  vl_encode_y(field = "Year:O") %>%
  vl_scale_x(paddingInner = 0) %>%
  vl_mark_rect()

text <- vl_chart() %>%
  vl_encode_color(value = "white") %>%
  vl_encode_x(field = "Cylinders:O") %>%
  vl_encode_y(field = "Year:O") %>%
  vl_encode_text(aggregate = "count", type = "Q") %>%
  vl_mark_text(baseline = "middle")

vl_layer(heatmap, text) %>%
  vl_add_data_url("https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_facet_column(field = "Origin", type = "nominal") 
```

### Repeat and concatenation {.tabset}

#### Horizontal concatenation

```{r}
bar_chart <- vl_chart() %>%
  vl_mark_bar() %>%
  vl_encode_x(field = "date:O", timeUnit = "month") %>%
  vl_encode_y(field = "precipitation:Q", aggregate = "mean")

point_chart <- vl_chart() %>%
  vl_mark_point() %>%
  vl_encode_x(field = "temp_min:Q", bin = TRUE) %>%
  vl_encode_y(field = "temp_max:Q", bin = TRUE) %>%
  vl_encode_size(aggregate = "count", type = "quantitative")
  
vl_hconcat(bar_chart, point_chart) %>% 
  vl_add_data(url = "https://vega.github.io/vega-lite/data/weather.csv") %>%
  vl_filter("datum.location === 'Seattle'")
```

#### Repeat wrap

```{r repeat-wrap}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_mark_bar() %>%
  vl_encode_x(field = list(`repeat` = "repeat"), type = "quantitative", bin = TRUE) %>%
  vl_encode_y(aggregate = "count", type = "quantitative") %>%
  vl_encode_color(field = "Origin", type = "nominal") %>%
  vl_repeat_wrap("Horsepower", "Miles_per_Gallon", "Acceleration", "Displacement")
```

#### Repeat color

```{r}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_mark_point() %>%
  vl_encode_x(field = "Horsepower", type = "quantitative") %>%
  vl_encode_y(field = "Miles_per_Gallon", type = "quantitative") %>%
  vl_encode_color(field = list("repeat" = "column"), type = "nominal") %>%
  vl_repeat_col("Origin", "Cylinders") %>%
  vl_resolve_scale_color("independent")
```


### Interactive multiple-view {.tabset}

#### Overview and detail

Brush on the bottom plot to zoom in on the top plot.

```{r}
detail <- vl_chart(width = 480) %>%
  vl_mark_area() %>%
  vl_encode_x('date:T') %>%
  vl_scale_x(domain = list(param = 'brush')) %>%
  vl_axis_x(title = '') %>%
  vl_encode_y('price:Q')

overview <- vl_chart(width = 480, height = 60) %>%
  vl_mark_area() %>%
  vl_add_interval_selection(name = 'brush', encodings = list('x')) %>%
  vl_encode_x('date:T') %>%
  vl_encode_y('price:Q') %>%
  vl_axis_y(tickCount = 3, grid = FALSE)


vl_vconcat(detail, overview) %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/sp500.csv") 

```

#### Cross-filter scatterplot matrix

Select an area on any sub-plot to color points on all plots.

```{r repeat-cross-filter}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_mark_point() %>%
  vl_encode_x(field = list("repeat" = "column"), type = "quantitative") %>%
  vl_encode_y(field = list("repeat" = "row"), type = "quantitative") %>%
  vl_add_interval_selection("brush", resolve = "global") %>%
  vl_encode_color(value = "grey") %>%
  vl_condition_color(param = "brush", field = "Origin", type = "nominal") %>%
  vl_repeat_row("Horsepower", "Acceleration", "Miles_per_Gallon") %>%
  vl_repeat_col("Miles_per_Gallon", "Acceleration", "Horsepower")
```


