---
title: "How to make a plot using vlbuildr"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette covers the basics of how to make a chart using `vlbuildr`.For more about the design philosophy of the package, see the [design vignette](https://vegawidget.github.io/vlbuildr/articles/design.html). For example charts see the [examples vignette](https://vegawidget.github.io/vlbuildr/articles/articles/examples.html).

# The three basic components

All plots require **data**, a **mark** and **encodings**. To intialize a chart you use

```{r}
library(vlbuildr)
vl_chart()
```

Without the other components it will be empty! But this output can be piped into commands to add the other components. Some components can also be added directly via that function call.  

## data

```{r}
vl_chart() %>%
  vl_add_data(values = mtcars)
```
Still empty! Need to set mark and encoding

We could also have passed the data directly to `vl_chart`:

```{r}
vl_chart(mtcars) 
```

Data can also be passed as a url using `vl_add_data_url` or the `url` argument to `vl_add_data`. `vl_chart` should also be able to recognize if the argument to `data` is a url and treat it as such.

## Mark

The 'mark' is the type of shape being plotted, e.g. 'point', 'line', 'bar'.

```{r}
vl_chart() %>%
  vl_add_data(values = mtcars) %>%
  vl_mark_point()
```

Without encodings, all the points show up on top of each other!

It can be passed without arguments besides the input spec (which often will be piped in), but also can take in arguments to change its properties. 

```{r}
vl_chart() %>%
  vl_add_data(values = mtcars) %>%
  vl_mark_point(color = 'red')
```

## Encoding

The encoding tells the plot what to use along the x or y axis, as well as what to use for color, tooltips, links, and other aspects of the plot that might be controlled by a column of your data. Encodings can also take constant values (generally with the `value` argument).

Functions to add an encoding are of the format `vl_encode_*` where the `*` is the name of the encoding, e.g. `x`:

```{r}
vl_chart() %>%
  vl_add_data(values = mtcars) %>%
  vl_mark_point() %>%
  vl_encode_x(field = 'wt', type = 'quantitative')
```

Multiple encodings can be added; it does not matter what order they are added.

```{r}
vl_chart() %>%
  vl_add_data(values = mtcars) %>%
  vl_mark_point() %>%
  vl_encode_x(field = 'wt', type = 'quantitative') %>%
  vl_encode_y(field = 'mpg', type = 'quantitative')
```

### Type shorthands

A shorthand can be used to specify the type. The field name can be joined using a colon with a single letter abbreviation. The abbreviations are:


```{r}
vl_chart() %>%
  vl_add_data(values = mtcars) %>%
  vl_mark_point() %>%
  vl_encode_x('wt:Q') %>%
  vl_encode_y('mpg:Q')
```

### Type inference

When the data has been added to the spec already and is given directly by an R data.frame (rather than a URL or a transformation), the type can also be inferred and does not need to be given.

```{r}
vl_chart() %>%
  vl_add_data(values = mtcars) %>%
  vl_mark_point() %>%
  vl_encode_x('wt') %>%
  vl_encode_y('mpg')
```

### `vl_encode`

Each encoding can be passed by a separate function, or a single call to `vl_encode` can be used:

```{r}
vl_chart() %>%
  vl_add_data(values = mtcars) %>%
  vl_mark_point() %>%
  vl_encode(x = 'wt', y = "mpg")
```

When not providing additional arguments for x or y encoding this form can be more succinct, but specifying each encoding separately can be easier when passing lots of arguments.  

For help constructing a full encoding object to `vl_encode` one can use functions within `vl` list of functions.

```{r}
vl_chart() %>%
  vl_add_data(values = mtcars) %>%
  vl_mark_point() %>%
  vl_encode(
    x = vl$X(field = 'wt', title = "weight"), 
     y = "mpg"
    )
```

# Interactivity

## Tooltips

There are three ways to get tooltips:

### All the encodings

To get a tooltip that shows the values for fields set in the encoding, use `tooltip = TRUE` in the call to add a mark to the plot.

```{r}
vl_chart(mtcars) %>%
  vl_mark_point(tooltip = TRUE) %>%
  vl_encode_x('wt') %>%
  vl_encode_y('mpg')
```

### All data associated with point

To get all the data to show up (not just what is used by encoding) pass `list(content = "data")` to the tooltip argument.

```{r}
vl_chart(mtcars) %>%
  vl_mark_point(tooltip = list(content = "data")) %>%
  vl_encode_x('wt') %>%
  vl_encode_y('mpg')
```

`vlbuildr` also has a shortcut, whereby if you pass "data" or "encoding" to the tooltip argument, it will transform it into the above list format in the spec. So the following will do the same thing:

```{r}
vl_chart(mtcars) %>%
  vl_mark_point(tooltip = "data") %>%
  vl_encode_x('wt') %>%
  vl_encode_y('mpg')
```

### To some constant value 

You can also set the tooltip to a constant value -- e.g. to get each point to say "Hi!":

```{r}
vl_chart(mtcars) %>%
  vl_mark_point(tooltip = "Hi!") %>%
  vl_encode_x('wt') %>%
  vl_encode_y('mpg')
```

### Via an encoding channel

We can also use a column of the data by passing the tooltip as an encoding. If we want to use the `row.names` of the `data.frame` we can use `'_row'` as the name:    

```{r}
vl_chart(mtcars) %>%
  vl_mark_point() %>%
  vl_encode_x('wt') %>%
  vl_encode_y('mpg') %>% 
  vl_encode_tooltip('_row')
```

## Parameters

More to come here... 

## Selections

More to come here... 

# View Composition

Sub-plots or 'views' can be combined in a number of ways. Layering assembles views into the same plot. Facetting splits views based on a column of data with the same encodings per view. Repeating splits views based on what field to use as one or two of the encodings. Concatenation combines arbitrary views together.  

## Layering

Layering assembles views into the same plot. A single view can only have one 'mark', so layering is how a plot can be made with multiple marks. Those marks can have different encodings or the same encodings.

### vl_layer

Two charts can be layered via `vl_layer`. Additional modifications can then be made to the resulting layered chart, although there are some limitations to the kinds of modifications that can be made to layered charts versus single view charts.  

```{r}
points_plot <- vl_chart() %>%
  vl_encode_x(field = "IMDB_Rating", type = "quantitative") %>%
  vl_encode_y(type = "quantitative", aggregate = "count") %>%
  vl_bin_x(maxbins=10) %>%
  vl_mark_bar()
  
avg_plot <- vl_chart() %>%
  vl_encode_x(field = "IMDB_Rating", type = "quantitative", aggregate = "mean") %>%
  vl_encode_color(value = "red") %>%
  vl_encode_size(value = 5) %>%
  vl_mark_rule()

vl_layer(points_plot, avg_plot) %>%
  vl_add_data_url("https://vega.github.io/vega-editor/app/data/movies.json")
```

In this case the data is added afterwards. That can help make sure the data isn't duplicated, and helps keep the size of the spec smaller. Note that layers can have different data, too, in which case the data would get added to each chart first before layering.

### The `+` operator

Layering can also be performed by using the `+` operator.

```{r}
(points_plot + avg_plot) %>%
  vl_add_data_url("https://vega.github.io/vega-editor/app/data/movies.json")
```

## Facet

Facetting splits up your data into different subplots based on a field in the data. This can be done across rows and/or columns.  Facets can get added in two ways: via `vl_facet_` functions or as encodings, eg `vl_encode_row`.

In terms of which option to use, they have slightly different options in terms of the parameters that can be given, but often are interchangeable.

### Facets as encoding

This example splits the data between rows based on the country of origin.

```{r}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_mark_bar() %>%
  vl_encode_x(field = "Horsepower", type = "quantitative") %>%
  vl_bin_x(maxbins = 15) %>%
  vl_encode_y(aggregate = "count", type = "quantitative") %>%
  vl_encode_row(field = "Origin", type = "nominal")
```

### Adding facets with facet functions

The following does the same as above, except via the `vl_facet_row` function.

```{r}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_mark_bar() %>%
  vl_encode_x(field = "Horsepower", type = "quantitative") %>%
  vl_bin_x(maxbins = 15) %>%
  vl_encode_y(aggregate = "count", type = "quantitative") %>%
  vl_facet_row(field = "Origin", type = "nominal")
```

## Repeating

Adding a 'repeat' is similar in some ways to facetting, except the same data gets shown in each subplot. Instead, what changes between subplots is what field gets used in the encoding. 

To use 'repeat' one needs to both call a `vl_repeat_*` function (where `*` is `wrap`, `row`, or `column`) with the field names to use, and specify the `field` for an encoding as `repeat:*` with the matching value of `*`.For example:

```{r}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_mark_bar() %>%
  vl_encode_x(field = "repeat:wrap", type = "quantitative", bin = TRUE) %>%
  vl_encode_y(aggregate = "count", type = "quantitative") %>%
  vl_encode_color(field = "Origin", type = "nominal") %>%
  vl_repeat_wrap("Horsepower", "Miles_per_Gallon", "Acceleration", "Displacement")
```

Alternatively a longer specification for the field can be used:

```{r}
vl_chart() %>%
  vl_add_data(url = "https://vega.github.io/vega-editor/app/data/cars.json") %>%
  vl_mark_bar() %>%
  vl_encode_x(field = list(`repeat` = "repeat"), type = "quantitative", bin = TRUE) %>%
  vl_encode_y(aggregate = "count", type = "quantitative") %>%
  vl_encode_color(field = "Origin", type = "nominal") %>%
  vl_repeat_wrap("Horsepower", "Miles_per_Gallon", "Acceleration", "Displacement")
```

## Concatenation

Different charts can be concatenated together via `vl_concat`, `vl_hconcat`, or `vl_vconcat` to concatenate in wrapping fashion (with number of columns set by `columns` arguments), horizontally, or vertically.  Data can be set for each individual view (and can then be different) or can be set for the concatenated version.

```{r}
bar_chart <- vl_chart() %>%
  vl_mark_bar() %>%
  vl_encode_x(field = "date:O", timeUnit = "month") %>%
  vl_encode_y(field = "precipitation:Q", aggregate = "mean")

point_chart <- vl_chart() %>%
  vl_mark_point() %>%
  vl_encode_x(field = "temp_min:Q", bin = TRUE) %>%
  vl_encode_y(field = "temp_max:Q", bin = TRUE) %>%
  vl_encode_size(aggregate = "count", type = "quantitative")
  
vl_hconcat(bar_chart, point_chart) %>% 
  vl_add_data(url = "https://vega.github.io/vega-lite/data/weather.csv") %>%
  vl_filter("datum.location === 'Seattle'")
```

Concatenation can also be done via the `|` or `&` operators, to do horizontal or vertical concatenation respectively.  

```{r}
(bar_chart | point_chart) %>% 
  vl_add_data(url = "https://vega.github.io/vega-lite/data/weather.csv") %>%
  vl_filter("datum.location === 'Seattle'")
```

```{r}
(bar_chart & point_chart) %>% 
  vl_add_data(url = "https://vega.github.io/vega-lite/data/weather.csv") %>%
  vl_filter("datum.location === 'Seattle'")
```
